GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers

### 2015-04-05 doc/report/midterm/article.tex
added more stuff to data center profiling

%%\documentclass[]{spie}  %>>> use for US letter paper
\documentclass[a4papert,12pt]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage[sort, numbers]{natbib}
\usepackage{hyperref}
\usepackage{adjustbox,lipsum}
\graphicspath{ {/Users/anirudhnair/code/GeoSched/src/GoogleTrace/plots} }
\usepackage{subfig}
\usepackage[margin=0.5in]{geometry}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Center Profiling } 
Google data-centers show 50\% less energy consumption than a typical data center. Google is able to achieve a trailing twelve-month (TTM) PUE of 1.12 across all data centers, in all seasons, including all sources of overhead. These data-centers have varied characteristics over time, which will be discussed in detail in this section.  
\subsection{Geo-Distributed Data Center}
We have chosen to simulate Google data centers that do the work for us around the world, around the clock at five different locations. These consists of two of them in United States namely Berkeley County, South Carolina and Lenoir, North Carolina. Other data centers that we have simulated are located one each in Europe, Asia, and South ?America. 
\subsection{Profile Parameters}
There are five techniques that Google typically uses to minimize its energy consumption and electricity costs. Frequent measuring of Power Usage Effectiveness is the key for determining the power usage for non-computing functions like cooling and power distribution. Secondly they manage airflow at data-centers, which allows cooling the hot spots available in data-centers. Moreover by setting thermostat to 80$^{\circ}$F it saves significant amount of electricity consumption and costs. Google researchers propose maximum usage of ambient temperature through evaporative cooling, cold reservoir, and airside economizer. This discards conventional chillers, which are the source of maximum energy consumption at data centers.  Altogether Google data continuously draw almost 260 million watts, which is about a quarter of the output of a nuclear power plant. Google uses renewable energy to power more than 35\% of its operations. These are wind, solar energy typically available at all data centers.

Google results show that data centers incur high PUE during summer and low PUE during winters. It clearly shows that surrounding temperature plays a critical role in deciding energy consumption at data centers. Data centers typically purchase wholesale electricity where the price is determined by day ahead market. 
\begin{figure} %
    \centering
    \subfloat[Daily Average Temperature ]{{\includegraphics[width=9cm]{AverageTemp} }}%
    \subfloat[Daily Maximum Variation]{{\includegraphics[width=9cm]{Variation} }}%
    \caption{Temperature variation for the year 2013}%
    \label{fig:temp}%
\end{figure}

\subsubsection{Temperature}
We use temperature as a key parameter to decide the location of job, as temperatures are diverse across data centers and encounters frequent fluctuations in temperature. In this case we make our assumption concrete by doing an empirical analysis of historical climate data, which provides substantial evidence. We extracted these historical climate data from various data repositories of the National Climate Data Center for all five locations covering the entire one-year period of 2013.

@@

\subsubsection{Electricity Pricing}
\subsubsection{Energy Source}

\subsection{Air Economizer}

@@

\section{Scheduling Algorithm}

\subsection{Features considered}
\subsubsection{Energy Cost Model}
%\subsubsection{Optimization Techiques}
%\subsubsection{Design}

>>>>

%%\documentclass[]{spie}  %>>> use for US letter paper
\documentclass[a4paper,12pt]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage[sort, numbers]{natbib}
\usepackage{hyperref}
\usepackage{adjustbox,lipsum}
\graphicspath{ {/Users/anirudhnair/code/GeoSched/src/GoogleTrace/plots} }
\usepackage{subfig}
\usepackage[margin=0.5in]{geometry}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Center Profiling } \label{sec:datacenter}
In order to run our experiments, we need to collect information about data centers that will help us make better scheduling decisions. We call this process data center profiling. The major energy costs in data center are from running the IT resources and from cooling the data centers. We create a profile for each of the data center with the information that effects the energy costs.
\subsection{Geo-Distributed Data Center}
Most cloud service providers including Google, Amazon, Facebook and Apple have data centers distributed across the globe. For our experiments, we consider five locations where Google has its data centers\cite{googledatacenter}. Table \ref{tab:datacenter} provides the location and labels for each of the data center.

\begin{table}[ht] \label{tab:datacenter}
\begin{tabular}{lll}
Location & Label & Associated Trace \\
Berkeley County, South Carolina & DC1 & A \\
Lenoir, North Carolina & DC2 & B \\
Singapore & DC3 & C \\
Quilicura, Chile & DC4 & D \\
Dublin, Ireland & DC5 & 
\end{tabular}
\end{table}

\subsection{Profile Parameters}
We consider two major parameters that has the most effect on the energy cost. They are a) Temperature b) Electricity. We collect temperature and electricity prices for these locations from historical archives. If time permits, we wish to also consider green energy available at the data center locations. We look at each of these parameters and how they effect energy costs in sections \ref{sec:temp}, \ref{sec:elec} and \ref{sec:green}.

\begin{figure} %
    \centering
    \subfloat[Daily Average Temperature ]{{\includegraphics[width=9cm]{AverageTemp} }}%
    \subfloat[Daily Maximum Variation]{{\includegraphics[width=9cm]{Variation} }}%
    \caption{Temperature variation for the year 2013}%
    \label{fig:temp}%
\end{figure}

\subsubsection{Temperature} {sec:temp}
Data centers located in cold regions have the advantage of using outside air to cool the data center. An air-side economizer brings outside air into a building and distributes it to the servers. Instead of being re-circulated and cooled, the exhaust air from the servers is simply directed outside. This can result in million dollars of saving annually. A difference in temperature across data centers mean different calling costs at these data centers. We wish to utilize this difference by scheduling jobs in low temperature data centers. In this case we make our assumption concrete by doing an empirical analysis of historical climate data. We extracted these historical climate data from various data repositories of the National Climate Data Center \cite{cdc} for all five locations covering the entire one-year period of 2013.

@@

\subsubsection{Electricity Pricing} {sec:elec}
Creation of power trading zones have resulted in wholesale power purchase under dynamic pricing schemes. Such markets are popular in USA \cite{elecusa}, Germany \cite{elecgermany} and other countries. Based on supply and demand bids these there can be high fluctuations in prices \cite{benini2002day}. There are many ways to predict there prices \cite{contreras2003arima} \cite{garcia2005garch}. If we can predict the electricity prices for the job duration, then we will be able to schedule the job such that the running costs go down.

\subsubsection{Energy Source} {sec:green}
Solar and wind energy are available in most of the modern data centers. Similar to electricity price predictions green energy prediction can also be done \cite{goiri2012greenhadoop} \cite{aksanli2012utilizing}. Since, this part is in exploratory stage, we will provide more information during the final report.

\subsubsection{Case Study: Google Data Centers} {sec:CSgoogle}
Google data-centers show 50\% less energy consumption than a typical data center. Google is able to achieve a trailing twelve-month (TTM) PUE of 1.12 across all data centers, in all seasons, including all sources of overhead.
There are five techniques that Google typically uses to minimize its energy consumption and electricity costs. Frequent measuring of Power Usage Effectiveness is the key for determining the power usage for non-computing functions like cooling and power distribution. Secondly they manage airflow at data-centers, which allows cooling the hot spots available in data-centers. Moreover by setting thermostat to 80$^{\circ}$F it saves significant amount of electricity consumption and costs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
